---
title: "p8105_hw1_yq2279"
author: "Qi Yuchen"
date: 2019-09-15
output: github_document
---

# Problem 1

```{r setup 1}
library(tidyverse)
```

First, we create a dataframe containing variables of four different types, and try to take the mean of each variable in the dataframe.

```{r dataframe and mean}
# create a dataframe
df = tibble(
  vec_numeric = rnorm(8),
  vec_logical = vec_numeric > 0,
  vec_char = c("My", "name", "is", "Chloe", ".", "Good", "day", "."),
  vec_factor = factor(c("male", "male", "female", "female", "trans", "trans", "male", "female"))
)

# take the mean of each variable
mean(pull(df, vec_numeric))
mean(pull(df, vec_logical))
mean(pull(df, vec_char))
mean(pull(df, vec_factor))
```

From the above we can see when taking the mean, the numerical vector and logical vector work, the character vector and the factor vector don't.

Next, we convert the logical, character, and factor variables to numeric.

```{r convert to mumeric, results = "hide", warning = FALSE}
as.numeric(pull(df, vec_logical))
as.numeric(pull(df, vec_char))
as.numeric(pull(df, vec_factor))
```

The logical variable is converted to a numeric varible with elements "0" and "1". This is because "0" means "false" and "1" means "true", which also explains why logical varibles can take the mean.
The factor variable is converted to a numeric varible with elements "1", "2" and "3". This is because the elements in the fator variable have three levels. But you cannot take the mean because there is no mean for levels.
The character variable is converted to a numeric varible with elements NA. This is because different character elements cannot be assigned a numeric value. So it cannot take the mean.


Now, we do the last part of problem 1.

Convert the logical vector to numeric, and multiply the random sample by the result.

```{r}
# convert the logical vector to numeric
logical_to_num = as.numeric(pull(df, vec_logical))

# multiply the random sample by the result
logical_to_num_multi_s = logical_to_num * pull(df, vec_numeric)
```

Convert the logical vector to a factor, and multiply the random sample by the result.

```{r}
# convert the logical vector to a factor
logical_to_fac = as.factor(pull(df, vec_logical))

# multiply the random sample by the result
logical_to_fac_multi_s = logical_to_fac * pull(df, vec_numeric)
```

Convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result.

```{r}
# convert the logical vector to a factor
logical_to_fac = as.factor(pull(df, vec_logical))

#  convert the result to numeric
logical_to_fac_to_num = as.numeric(logical_to_fac)

# multiply the random sample by the result
logical_to_fac_to_num_multi_s = logical_to_fac_to_num * pull(df, vec_numeric)
```

# Problem 2

```{r setup 2}
rm(list = ls())
library(tidyverse)
```

First we Create a data frame.

```{r dataframe}
df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical = x + y > 1,
  vec_num = as.numeric(vec_logical),
  vec_fac = as.factor(vec_logical)
)
```

* The size of the dataset is nrow = `r dim(df)[1]`, ncol = `r dim(df)[2]`. 

*  the mean, median, and standard deviation of x is `r mean(pull(df, x))`, `r median(pull(df, x))`, and `r sd(pull(df, x))`.

Then make a scatterplot of y vs x; color points using the logical variable.

```{r plot logical variable}
ggplot(df, aes(x = x, y = y, color = vec_logical)) + geom_point()

ggsave(file = "plot_logical_variable.png")
```

Color points using the numeric variable.

```{r plot numeric variable}
ggplot(df, aes(x = x, y = y, color = vec_num)) + geom_point()
```

Color points using the numeric variable.

```{r plot factor variable}
ggplot(df, aes(x = x, y = y, color = vec_fac)) + geom_point()
```
